---
title: "The trackdown Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{trackdown-workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  eval = FALSE,
  comment = "#>"
)
```

## The Workflow Idea

When collaborating on the writing of a `.Rmd` (or `.Rnw`) document it is important to consider separately code and prose:

- **Code** - Collaborative code writing is done efficiently following traditional **Git** workflow based on an online repository (e.g., GitHub or GitLab)
- **Prose** - Collaborative prose writing is done efficiently on **Google Docs** where the familiar and simple online interface allows multiple users to simultaneously write and edit the same document

Thus, the workflow main idea is very simple: upload the `.Rmd` (or `.Rnw`) document to Google Drive for collaborative prose writing in Google Docs and download the document locally to continue working on the code using Git. This iterative process of uploading to and downloading from Google Drive continues until the desired results are obtained. The workflow can be summarized as:

> Collaborative **code** writing using **Git** and Collaborative **prose** writing on **Google Docs** 

The package `trackdown` offers different functions to manage this workflow:

- `upload_file()` - upload a file for the first time to Google Drive
- `update_file()` - update the content of an existing file in Google Drive with the contents of a local file
- `download_file()` - download edited version of a file from Google Drive updating the local version 
- `render_file()` - download and render a file from Google Drive

## Workflow Example

Here a hypothetical workflow using `trackdown` is described as an example. The actual workflow depends on specific needs and conditions.

Suppose you want to collaborate with your colleagues on the writing of an R Markdown document. If you are the most experienced among your colleagues with R and in programming in general, it would be better if you manage and organize the workflow. 

#### Upload File

You create the initial document, for example `My-Report.Rmd`, and upload the file in Google Drive using the function `upload_file()`:

```{r, echo = TRUE, eval = FALSE}
trackdown::update_file(file = "Project-A/My-Report.Rmd")
```

In this way, the `My-Report.Rmd` file is uploaded from your local computer to your Google Drive. By default, the file is uploaded in Google Drive using the same name as the local file (without extension) in a folder named `tackdown/`. 

Now you can share the link to the document in your Google Drive with your colleagues and invite them to collaborate on the writing of the prose, using the familiar Google Docs interface. Note that working on the code part of the document should be avoided as it is not possible to run any code in Google Drive.

#### Download File

At some point, you will want to add some code to include figures, tables or results from the analyses. This must not be done directly in Google Docs, so first you are required to download the document.

Before downloading the document from Google Drive, however, it is important that you accept all changes made to the document in Google Docs. You can simply select from the top menu “*Tools* > *Review suggested edits* > *Accept all*”. Now the edited version of the document can be downloaded from Google Drive using the function `download_file()`:

```{r, echo = TRUE, eval = FALSE}
trackdown::download_file(file = "Project-A/My-Report.Rmd")
```

Note that downloading the file from Google Drive will overwrite the local file.  

#### Update File

Once you added the required code chunks, further editing on the prose may still be necessary. In this case, you first update the file in Google Drive with your local version of the document using the function `update_file()`:

```{r, echo = TRUE, eval = FALSE}
trackdown::update_file(file = "Project-A/My-Report.Rmd")
```

In this way, the document in Google Drive is updated with your latest changes. Now you and your colleagues can continue collaborating on the writing of the document. Note that updating the file in Google Drive will overwrite its current content losing tracked changes.


#### Render File

Once the document is finished and all changes have been accepted, you can download the document and automatically compile the document to obtain the actual output using the function `render_file()`:

```{r, echo = TRUE, eval = FALSE}
trackdown::render_file(file = "Project-A/My-Report.Rmd")
```

This function is simply a wrapper around `download_file()` that will call `rmarkdown::render()` once the document has been downloaded.

#### A Quick Check

Taking advantage of the "*Undo*" option available in Google Docs, it is possible to download the current document version at any time during the workflow without the need to accept permanently all suggested edits. You can:

1. Select in Google Docs from the top menu “*Tools* > *Review suggested edits* > *Accept all*” to momentarily accept all changes
2. Download the current version of the document using `download_file()` (or `render_file()`) function
3. In Google Drive press "*Undo*" to restore the document to the previous state (i.e., before accepting all changes)

In this way it is possible to check the current document and compile it locally at any time, without the need to accept permanently all suggested edits. Note, however, that is not possible to update the document in Google Drive with some local changes. This would overwrite the current content losing tracked changes.  

## Collaboration: Prose vs Code

As stated above, when collaborating on the writing of a `.Rmd` (or `.Rnw`) document it is important to consider separately code and prose. Collaborative code writing should be based on version control systems like Git, whereas collaborative prose writing is done more efficiently in Google Docs. Let's further discuss these two aspects:

- **Code** - Collaborative code writing is better managed using version control systems like **Git**. Ideally, collaborators working on the code are experienced enough to organize the project in an online repository (e.g., GitHub or GitLab) and manage the workflow using Git. Writing or editing actual code on Google Docs, although possible, is strongly discouraged as it is no possible to run the code to check if it is correct. Thus, it is very easy to make several mistakes that would break the code. It is much better to write code in an appropriate IDE  (e.g., RStudio).

- **Prose** - To facilitates the collaborative prose writing `trackdown` moves the whole process on **Google Docs**. Collaborators, however, should be aware that when the document is downloaded from Google Drive it is still considered as a `.Rmd` (or `.Rnw`) file. Thus, any formatted text (e.g., bold, italic, titles, font size, etc.) made in Google Docs will be lost. To effectively maintain text formatting options, they have to be specified using proper Markdown (or LaTeX) syntax. Collaborators with no experience in programming could find this too demanding. In this case, they might prefer to modify the text format using the common interface commands, leaving to their more experienced collaborators the task of *translating* the text formatting options into the appropriate Markdown (or LaTeX) commands.

Overall, the workflow proposed by `trackdown` for collaborative writing of an R Markdown (or Sweave) document is an iterative process where the document is uploaded/updated to Google Drive for collaborative prose writing in Google Docs and downloaded locally for collaborative code writing using Git. 

Note that the main limitations of this workflow are that it is not possible to simultaneously collaborate on the prose and the code included in the document. In this case, changes would be made on two different copies of the same document. Changes on the prose would be saved on the Google Drive copy of the document, whereas changes on the code would be saved on the local copy of the document. These two document versions can not be merged together but only overwritten, losing one of the two versions. However, minimal organization of the workflow into sequential steps will guarantee a smooth experience.
