---
title: 'Debugging Issue #17'
author: "trackdown"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if(!requireNamespace("trackdown")) stop("trackdown is required, plase install trackdown first")
```
## Isssue
https://github.com/ClaudioZandonella/trackdown/issues/17

### Uploading

Identifying header code when uploading

```{r}
# !! add here the name of your Rmd document 
name_your_file <- "YourReport.Rmd"  #  <---- !!change here!!
file_info <- trackdown:::get_file_info(name_your_file)
original_text <- readLines(name_your_file)
```

Check patterns used to retrieve document header and code chunks.

```{r}
info_patterns <- trackdown:::get_extension_patterns(extension = file_info$extension)
info_patterns
```

Extract document header
```{r}
trackdown:::extract_header(original_text, info_patterns)
```

Check if header is correctly substituted

```{r}
# print first lines of the text to upload
text_to_upload <- trackdown:::hide_code(original_text, file_info)[1:7]
text_to_upload
```

Check text ready to upload

```{r}
# Print first 600 charcters to evalaute [[document-header]] absence/presence
text_to_upload <- trackdown:::format_document(text_to_upload, file_info, hide_code = TRUE)
substr(text_to_upload, start = 1, stop = 1000)
```

Cat document to your local compute
```{r}
cat(text_to_upload, file = "document-to-upload.txt")
```

a file named `document-to-upload.txt` will be created in the same folder as the current report. This file contains the text that will be uploaded to Google drive. Check if [[document-header]] tag is present

```{r}
readLines("document-to-upload.txt")[1:13]
```

remove created file `document-to-upload.txt`

```{r}
unlink("document-to-upload.txt")
```

